---
title: "Activity prediction using machine learning"
author: "Shashikant Singh Kunwar"
date: "13/09/2021"
output:
  slidy_presentation: default
  ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```



## Introduction
### AIM
* Using machine learning method to predict physical activities.

### UTILITY
* This can help users get better and instant feedback on the efficiency of their exercises.
* Allows users to easily track their performance over period of time.
* Users can have much better control over their activities and plan workout accordingly.

### HOW TO ACHIEVE IT
* Users can monitor their physical activities through sensors on wearable devices like
watches, bands. 
* These sensors can be attached to different body parts/clothing based on the exercise 
 or on instruments users are using for exercise.
* These sensors can collect data and the computation can be performed on either the wearable device
or on mobile phone these sensors are connected to.


## EXAMPLE: HAR Dataset
* Ugulino, W.; Cardador, D.; Vega, K.; Velloso, E.; Milidiu, R.; Fuks, H. **Wearable Computing: Accelerometers' Data Classification of Body Postures and Movements**. <a href="http://groupware.les.inf.puc-rio.br/work.jsf?p1=10335#ixzz4R3exXJyF"> Read more here </a>.
* Sensor data is collected for 6 users performing different activities.
* Machine learning model learns to predict the activity based on sensor data.
* Output is the way activity is performed.

```{r,echo=TRUE,cache=TRUE,warning=FALSE,message=FALSE}
library(caret)
# loading data
data<-read.csv("pml-training.csv")
# loading pre trained model
model<-readRDS("model2.rda")
# names of all the predictors
all_predictors<-names(data)

# name of columns that contains data for accelerometer
accel_predictors<-all_predictors[grep("*accel*",all_predictors)]
listna<-c()
for (j in seq(1:20)){
    if (sum(is.na(data[,accel_predictors[j]]))){
    listna<-c(listna,j)}}
  
# list of variable that don't have NA.
accel_nona_predictors<-accel_predictors[-listna]

# list with predictor names and classe
data_nona_columns<-c(accel_nona_predictors,"classe")
        
# return only finite data for accelerometer and labels
data<-data[,data_nona_columns]    

# set seed for reproducibility
set.seed(42)

# splitting the data
inTrain<-createDataPartition(data$classe,p=0.7,list=FALSE)
# using data that was not used for training
validation_data<-data[-inTrain,]

val_data<-data[-inTrain,]
val_data$classe<-as.factor(val_data$classe)
```


## Activity prediction
* The pre trained  hyper parameter optimized model can be loaded directly to save time 
and resources.
* Calculating confusion matrix for the val data (not used in training the model).
* Trained model can predict the activities correctly with $\approx 95\%$ accuracy.
```{r,echo=TRUE,cache=TRUE,message=FALSE,warning=FALSE}
library(caret)
set.seed(42)
# predicting labels of the validation data
preds<-predict(model,val_data[,-dim(val_data)[2]])
# calculating confusion matrix between predictions and labels
confusionMatrix(val_data$classe,preds)
```


## Visualization
* Plotting predictors(sensor data) with highest variable importance during training. 
* The color of each point represents the activity.
* For plotting we are using the name of activity and not the label used for machine learning.
```{r,cache=TRUE,echo=FALSE,warning=FALSE,message=FALSE}
library(hash)
library(rattle)
activity<-hash()
activity["A"]<- "Dumbbell Biceps Curl Done properly"
activity["B"]<- "Throwing elbows to front during Dumbbell Biceps Curl" 
activity["C"]<- "Lifting dumbbell only halfway during Dumbbell Biceps Curl"
activity["D"]<- "Lowering dumbbell only halfway during Dumbbell Biceps Curl"
activity["E"]<- "Throwing hips in front during Dumbbell Biceps Curl"
for (i in keys(activity)){
    validation_data[validation_data$classe==i,]$classe <- activity[[i]]
}
```

```{r,cache=TRUE,warning=FALSE,echo=TRUE,message=FALSE}
library(plotly)
fig<-plot_ly(x=~accel_belt_z,y=~accel_dumbbell_z,z=~accel_dumbbell_y,
        color=~classe,colors= c("red","blue","cyan","magenta","green"),
        data=validation_data,type="scatter3d",mode="markers")

fig<- layout(fig,scene=list(xaxis = list(title='accel_belt_z'),
        yaxis = list(title='accel_dumbbell_z'),
        zaxis=list(title='accel_dumbbell_y')))

fig
```

### Conclusion
* Machine learning can help users achieve their fitness goals.
* Instant feedback can immensely help the users.
* These models can make predictions using processing power of mobile phone or smart watch.
